Random Forest to find the top important features

### 2019-11-01 scripts/15_RF_find_top_features.Rmd
update

---
title: "Random Forest to find the top important features"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
```

@@

@@

There are much for Y patients than N. If all assigning to Y it will probably like this.

### Use all features + all pheno var as predictors

```{r} 
# bind the pheno with the transposed cts matrix
combined <- cbind(s_tb_rf, pheno %>% 
                    select(-Sampleid))

@@

### Use more abundant features(species abundance > 1%) + all pheno var as predictors

@@

combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
                    select(-Sampleid))

combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T )

@@

No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.

@@

Using the third table predictors.

@@

What's interesting is this species *Ruminococcus gnavus*. [This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines, which could be hypothesized to cause toxicity too.

>>>>

---
title: "Random Forest to find the top important features"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
library(arrangements)
```

@@

```{r}
# check whether the pheno vars are correlated (in pairs)
all_var <- names(pheno)
combi <- combinations(all_var[all_var != 'Sampleid'], 2) %>% 
  as.data.frame(stringsAsFactors = F)

phe_cor <- apply(combi, 1, function(Row) {
  phe_fil = pheno %>% 
    select(Row[['V1']], Row[['V2']]) %>% 
    table()

  chisq.test(phe_fil)$p.value
}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
  select(-V1, -V2) %>% 
  rename(corrected_pval = names(.)[1]) %>% 
  arrange(corrected_pval) %>% 
  mutate(CORR = if_else(corrected_pval < 0.1, 'Yes', 'No'))

phe_cor %>% 
  kable(caption = 'Test on independence between two of the pheno vars') %>% 
  kable_styling(full_width = F, position = 'left')
```

@@

Basically assigning everything to Y.

### Use all features + uncorrelated pheno var as predictors

Uncorrelated pheno var: ORR, Disease_Status_CAR for Toxicity

```{r} 
# bind the pheno with the transposed cts matrix
combined <- cbind(s_tb_rf, pheno %>% 
                    select(Toxicity, ORR, Disease_Status_CAR))

@@

### Use more abundant features(species abundance > 1%) + uncorrelated pheno var as predictors

@@

combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
                    select(Toxicity, ORR, Disease_Status_CAR))

combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = 1000 )

@@

No improvements at all. I think because the relatively small sample size and the signal itself is weak. There's no guarantee that there is association between microbiome and the outcome to begin with.

```{r}
# to use the top N percent features

# to test on different threshold of selecting most abundant species
sum_thre <- seq(0.5, 1, 0.05)
tree_num <- 1000

res <- sum_thre %>% 
  set_names(sum_thre) %>% 
  map(function(thre) {
    s_sum_top <- quantile(s_sum, thre)
    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
                    select(Toxicity, ORR, Disease_Status_CAR))

    tibble(
      num_fea = length(s_sum[s_sum > s_sum_top]) + 2,
      OOB_err_rate = randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]

    ) 
  }) %>% 
  bind_rows(.id = 'thre')

res %>% 
  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
  kable_styling(full_width = F, position = 'left')
```

@@

Using the table with features abundance > 1% and the two uncorrelated vars.

@@

What's interesting is this species *Ruminococcus gnavus*. [This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines.

And [this paper](https://www.ncbi.nlm.nih.gov/pubmed/29330424) mentions that "*Faecalibacterium prausnitzii* decrease and *Ruminococcus gnavus* increase were the essential characteristics in CHF patients' gut microbiota".
